### 2016-11-22 local-capability-machine.tex
Added size to malloc instruction

\begin{description}
\item[\texttt{call} $r_f(\bar{r}_{\var{args}},\bar{r}_{\var{priv}})$] $\bar{r}_{\var{args}}$ and $\bar{r}_{\var{priv}}$ are lists of registers. An overview of this call:
  \begin{itemize}
  \item Set up activation record
  \item Create local enter capability for activation
  \item Clear unused registers
  \item Jump
  \item Upon return: Run activation code
  \end{itemize}
A more detailed description of each of the above steps:
\begin{description}
  \item [Set up activation record]
    \begin{itemize}
    \item Run malloc to get a piece of memory with space for:
    \begin{itemize}
      \item Words in $\bar{r}_{\var{priv}}$
      \item Code return capability (opc)
      \item Activation code
    \end{itemize}
    \item Store the words in $\bar{r}_{\var{priv}}$ to the activation record. 
    \item Adjust the current pc to point to the return address and save it to the activation record.
    \item Write the activation code to the activation record.
  \end{itemize}
  \item [Create local enter capability for activation] Make the capability for the activation record point to the beginning of the activation record and restrict it to a local-enter capability. Place this capability in $r_0$.
  \item [Clear unused registers]
    Clear all the register that are not $\pcreg$, $r_0$, $r_f$ or in $\bar{r}_{\var{args}}$.
  \item [Jump] jump to $r_f$
  \item [Activation code] The activation code does the following:
    \begin{itemize}
    \item Move the stored ``private'' words in to their respective $\bar{r}_{\var{priv}}$ registers.
    \item Clear all registers but $r_1$, $\pcreg$, and $\bar{r}_{\var{priv}}$.
    \item Load the return address to $\pcreg$
    \end{itemize}
\end{description}
\item[\texttt{malloc x}] Calls malloc to allocates a piece of memory (of an appropriate size). In the program $\texttt{x}$ will henceforth be used to refer to the capability returned by malloc.
\item[\texttt{hstore x n}] Fetches the capability referred to with \texttt{x} and stores \texttt{n} to it.
\item[\texttt{assert x y}] Fetches \texttt{x} and \texttt{y} and compares them. If they are equal, then execution continues. If they are unequal, then the assertion flag is set to 1 and execution halts.
\end{description}

\subsection{Stack}
Some programs will assume access to a stack which will be in part indicated by the program macros, but also in the correctness lemma. The stack is accessed through a local $\rwlx$-capability. Programs will assume that the stack resides in some register, say $r_{\var{stk}}$.
\begin{description}
\item[\texttt{push r}] Pushes the word in register \texttt{r} to the stack by storing it through the stack capability and increment said capability.\footnote{The stack grows upwards.}
\item[\texttt{pop r}] Pops the top word of the stack by loading it to register $r$, moves 0 to the address, and decrementing the stack pointer.
\item[\texttt{scall} $r_f(\bar{r}_{\var{args}},\bar{r}_{\var{priv}})$] $\bar{r}_{\var{args}}$ and $\bar{r}_{\var{priv}}$ are lists of registers. This call assumes $r_{\var{stk}}$ contains a stack capability. An overview of this call:
  \begin{itemize}
  \item Push restore code to the stack.
  \item Push ``private'' registers to the stack.
  \item Push return address capability 
  \item Push stack pointer
  \item Create local enter capability for restoration
  \item Restrict stack capability
  \item Clear unused registers
  \item Jump
  \item Upon return: Run restore code
  \item Return address: Clear the stack segment we gave control over
  \item Restore ``private'' state
  \end{itemize}
A more detailed description of the above steps:
\begin{description}
  \item [Push restore code to the stack]
    Push the restore code to the stack (described later). This code needs to be on the stack to make sure the stack capability can be restored. We try to keep the restore code on the stack minimal.
  \item [Push ``private'' registers to the stack]
    Push all the words in the registers in  $\bar{r}_{\var{priv}}$ to the stack.
  \item [Push return address capability]
    Push a capability for the return address (on the heap) to the stack.
  \item [Push stack pointer]
    Push the full stackpointer to the heap.
  \item [Create local enter capability for restoration]
    Make a new version of the stack pointer that points to the beginning of the restoration code. Restrict it to a local-enter capability and put it in $r_0$.
  \item [Restrict stack capability]
    Make the stack capability only govern the unused part.
  \item [Clear unused registers]
    Clear all registers but $\pcreg$, $r_0$, $r_{\var{stk}}$, $r_f$, and $\bar{r}_{\var{args}}$.
  \item [Jump] Jump to $r_f$.
  \item [Restore code]
    Load the stack capability to $r_{\var{stk}}$. Pop the old program counter to $\pcreg$.
  \item [Return Address:] \forcenewline
    \begin{description}
    \item [Clear the stack segment we gave control over]
    \item [Restore ``private'' state] \forcenewline
      \begin{itemize}
      \item Pop the private state on the stack into their respective $\bar{r}_{\var{priv}}$ registers.
      \item Pop the restore code of the stack
      \end{itemize}
  \end{description}
\end{description}
\end{description}

\begin{figure}
  \label{fig:stack-before-call}
  \centering
  \begin{tabular}[!h]{r | >{\raggedright\arraybackslash}p{3cm} |}
\multicolumn{2}{l}{Stack} \\
   & \\
   & $\vdots$\\
\cline{2-2}
   & 0 \\
\cline{2-2}
$c_{\var{stk}} \rightarrow$   & local stack\\
   & $\vdots$
\end{tabular}
\hspace{1cm}
\begin{tabular}{r | >{\centering\arraybackslash}p{0.75cm} |}
\multicolumn{2}{r}{Register file} \\
\cline{2-2}
$\pcreg$ & $c_{\pc}$\\
\cline{2-2}
$r_0$  & $c_0$ \\
\cline{2-2}
$r_{\var{stk}}$  & $c_{\var{stk}}$ \\
\cline{2-2}
$r_f$ & $c_f$ \\
\cline{2-2}
$r_{\var{args},1}$ & $w_{a,1}$ \\
\cline{2-2}
& $\vdots$ \\
\cline{2-2}
$r_{\var{args},n}$ & $w_{a,1}$\\
\cline{2-2}
$r_{\var{priv},1}$ & $w_{p,1}$\\
\cline{2-2}
& $\vdots$ \\
\cline{2-2}
$r_{\var{priv},m}$ & $w_{p,m}$\\
\cline{2-2}
& $\vdots$ \\
\cline{2-2}
\end{tabular}
\caption{Stack and register-file before the call}
\end{figure}

>>>>

\begin{description}
\item[\texttt{call} $r_f(\bar{r}_{\var{args}},\bar{r}_{\var{priv}})$] $\bar{r}_{\var{args}}$ and $\bar{r}_{\var{priv}}$ are lists of registers. An overview of this call:
  \begin{itemize}
  \item Set up activation record
  \item Create local enter capability for activation
  \item Clear unused registers
  \item Jump
  \item Upon return: Run activation code
  \end{itemize}
A more detailed description of each of the above steps:
\begin{description}
  \item [Set up activation record]
    \begin{itemize}
    \item Run malloc to get a piece of memory with space for:
    \begin{itemize}
      \item Words in $\bar{r}_{\var{priv}}$
      \item Code return capability (opc)
      \item Activation code
    \end{itemize}
    \item Store the words in $\bar{r}_{\var{priv}}$ to the activation record. 
    \item Adjust the a copy of the current pc to point to the return address and save it to the activation record.
    \item Write the activation code to the activation record.
  \end{itemize}
  \item [Create local enter capability for activation] Make the capability for the activation record point to the beginning of the activation record and restrict it to a local-enter capability. Place this capability in $r_0$.
  \item [Clear unused registers]
    Clear all the register that are not $\pcreg$, $r_0$, $r_f$ or in $\bar{r}_{\var{args}}$.
  \item [Jump] jump to $r_f$
  \item [Activation code] The activation code does the following:
    \begin{itemize}
    \item Move the stored ``private'' words in to their respective $\bar{r}_{\var{priv}}$ registers.
%    \item Clear all registers but $r_1$, $\pcreg$, and $\bar{r}_{\var{priv}}$.
    \item Load the return address to $\pcreg$
    \end{itemize}
\end{description}
\item[\texttt{malloc $r$ $n$}] Calls malloc to allocates a piece of memory of size $n$. The capability will be stored in register \texttt{r}.
\item[\texttt{store $r$ $n$}] Stores $n$ to $r$ - notice the semantics do not allow an integer to be stored directly.
\item[\texttt{assert $r_1$ $r_2$}] Compares the words in register $r_1$ and $r_2$. If they are equal, then execution continues. If they are unequal, then the assertion flag is set to 1 and execution halts.
\end{description}
Note:
\begin{itemize}
\item \texttt{call} will fail if we have local capabilities as part of the ``private'' state as it relies on a capability returned by malloc which will not be permit-write-local. Below, we introduce \texttt{scall} which can handle local capabilities in the ``private'' state.
\end{itemize}
\subsection{Stack}
Some programs will assume access to a stack which will be in part indicated by the program macros, but also in the correctness lemma. The stack is accessed through a local $\rwlx$-capability. Programs will assume that the stack resides in some register, say $r_{\var{stk}}$.
\begin{description}
\item[\texttt{push $r$}] Pushes the word in register \texttt{r} to the stack by storing it through the stack capability and increment said capability.\footnote{The stack grows upwards.}
\item[\texttt{pop $r$}] Pops the top word of the stack by loading it to register $r$, moves 0 to the address, and decrementing the stack pointer.
\item[\texttt{scall} $r_f(\bar{r}_{\var{args}},\bar{r}_{\var{priv}})$] $\bar{r}_{\var{args}}$ and $\bar{r}_{\var{priv}}$ are lists of registers. This call assumes $r_{\var{stk}}$ contains a stack capability. An overview of this call:
  \begin{itemize}
  \item Push restore code to the stack.
  \item Push ``private'' registers to the stack.
  \item Push return address capability 
  \item Push stack pointer
  \item Create local enter capability for restoration
  \item Restrict stack capability
  \item Clear unused registers
  \item Jump
  \item Upon return: Run restore code
  \item Return address: Clear the stack segment we gave control over
  \item Restore ``private'' state
  \end{itemize}
A more detailed description of the above steps:
\begin{description}
  \item [Push restore code to the stack]
    Push the restore code to the stack (described later). This code needs to be on the stack to make sure the stack capability can be restored. We try to keep the restore code on the stack minimal.
  \item [Push ``private'' registers to the stack]
    Push all the words in the registers in  $\bar{r}_{\var{priv}}$ to the stack.
  \item [Push return address capability]
    Push a capability for the return address (on the heap) to the stack.
  \item [Push stack pointer]
    Push the full stackpointer to the stack.
  \item [Create local enter capability for restoration]
    Make a new version of the stack pointer that points to the beginning of the restoration code. Restrict it to a local-enter capability and put it in $r_0$.
  \item [Restrict stack capability]
    Make the stack capability only govern the unused part.
  \item [Clear unused registers]
    Clear all registers but $\pcreg$, $r_0$, $r_{\var{stk}}$, $r_f$, and $\bar{r}_{\var{args}}$.
  \item [Jump] Jump to $r_f$.
  \item [Restore code]
    Load the stack capability to $r_{\var{stk}}$. Pop the old program counter to $\pcreg$.
  \item [Return Address:] \forcenewline
    \begin{description}
    \item [Clear the stack segment we gave control over] Store 0 to all the cells the stack capability passed to the callee governed.
    \item [Restore ``private'' state] \forcenewline
      \begin{itemize}
      \item Pop the private state on the stack into their respective $\bar{r}_{\var{priv}}$ registers.
      \item Pop the restore code of the stack
      \end{itemize}
  \end{description}
\end{description}
\end{description}
Note:
\begin{itemize}
\item As a rule of thumb: If you have provided an untrusted entity access to part of the stack, then it needs to be cleared upon return.
\item As a rule of thumb: If you receive a stack from an untrusted source, then you need to check that it is a local $\rwlx$-capability and clear it! If any callbacks are provided, then they need to be global.
\end{itemize}
\begin{figure}
  \label{fig:stack-before-call}
  \centering
  \begin{tabular}[!h]{r | >{\raggedright\arraybackslash}p{3cm} |}
\multicolumn{2}{l}{Stack} \\
   & \\
   & $\vdots$\\
\cline{2-2}
   & 0 \\
\cline{2-2}
$c_{\var{stk}} \rightarrow$   & local stack\\
   & $\vdots$
\end{tabular}
\hspace{1cm}
\begin{tabular}{r | >{\centering\arraybackslash}p{0.75cm} |}
\multicolumn{2}{r}{Register file} \\
\cline{2-2}
$\pcreg$ & $c_{\pc}$\\
\cline{2-2}
$r_0$  & $c_0$ \\
\cline{2-2}
$r_{\var{stk}}$  & $c_{\var{stk}}$ \\
\cline{2-2}
$r_f$ & $c_f$ \\
\cline{2-2}
$r_{\var{args},1}$ & $w_{a,1}$ \\
\cline{2-2}
& $\vdots$ \\
\cline{2-2}
$r_{\var{args},n}$ & $w_{a,1}$\\
\cline{2-2}
$r_{\var{priv},1}$ & $w_{p,1}$\\
\cline{2-2}
& $\vdots$ \\
\cline{2-2}
$r_{\var{priv},m}$ & $w_{p,m}$\\
\cline{2-2}
& $\vdots$ \\
\cline{2-2}
\end{tabular}
\caption{Stack and register-file before the call}
\end{figure}
